Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-22 supplement.tex
Updates from ShareLaTeX
Add run times section to supplement

\newcommand{\ourprog}{\Software{RNARedPrint}}

@@

@@

\medskip
\begin{center}
\large Stefan Hammer, Yann Ponty, Wei Wang and Sebastian Will
\end{center}
\section{Approximate counting and random generation}
In fact, not only is $\#{\sf BIS}$ a reference problem in counting complexity, but it is also a landmark problem with respect to the complexity of approximate counting problems. In this context, it is the representative for a class of $\#{\sf BIS}$-hard problems~\citep{Bulatov2013} that are easier to approximate than $\# {\sf SAT}$, yet are widely believed not to admit any Fully Polynomial-time Randomized Approximation Scheme. Recent results reveal a surprising dichotomy in the behavior of $\#{\sf BIS}$: it admits a Fully Polynomial-Time Approximation Scheme (FPTAS) for graphs of max degree $\le 5$~\citep{Weitz2006}, but is as hard to approximate as the general $\#{\sf BIS}$ problem on graphs of degree $\ge 6$~\citep{Cai2016}. In other words, there is a clear threshold, in term of the max degree, separating (relatively) easy instances from really hard ones.

@@

>>>>

\newcommand{\RNAblueprint}{{\tt \bfseries{}\color{black!85} RNA\textcolor{blue!70!black}{Blue}Print}}
\newcommand{\ourprog}{{\tt \bfseries{}\color{black!85}RNA\textcolor{red!70!black}{Red}Print}}

@@

%% highlight changes for revision in red
\newcommand{\revised}[1]{{\color{red} #1}}
\newenvironment{revision}{\color{red}}{\color{black}}

@@

\medskip
\begin{center}
\large Stefan Hammer, Wei Wang, Sebastian Will and Yann Ponty
\end{center}
\section{Approximate counting and random generation}
In fact, not only is $\#{\sf BIS}$ a reference problem in counting complexity, but it is also a landmark problem with respect to the complexity of approximate counting problems. In this context, it is the representative for a class of $\#{\sf BIS}$-hard problems~\citep{Bulatov2013} that are easier to approximate than $\# {\sf SAT}$, yet are widely believed not to admit any Fully Polynomial-time Randomized Approximation Scheme. Recent results reveal a surprising dichotomy in the behavior of $\#{\sf BIS}$: it admits a Fully Polynomial-Time Approximation Scheme (FPTAS) for graphs of max degree $\le 5$~\citep{Weitz2006}, but is as hard to approximate as the general $\#{\sf BIS}$ problem on graphs of degree $\ge 6$~\citep{Cai2016}. In other words, there is a clear threshold, in term of the max degree, separating (relatively) easy instances from really hard ones.

@@

\begin{revision}
  \section{Run times for generating samples}

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times for generating 1000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{fig:run-times-redprint}).

\todo[inline]{@Stefan: could you please include technical data about the machines used for the timings? Please also check the claims/statements in the text for correctness. If necessary---please edit.}

A fair comparison to sampling times by other tools can hardly be given, since the other tools produce samplings of very different nature. In particular, Modena does not make any guarantees on properties of the sampling. Non-surprisingly, this allows it to sample significantly faster (several orders of magnitudes).

The comparison to run times of \RNAblueprint{} is still interesting, since this tool generates---similarly to our tool---\emph{controlled} samples. Nevertheless, as we argue the Boltzmann sampling is fundamentally more flexible than uniform sampling as generated by \RNAblueprint. Moreover, in this concrete case, \ourprog{} samples based on the rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width; here, the increase is exactly the same as the one from the base pair to the stacking model as reported in Figure~\ref{}. In this light, \ourprog{} performs even particularly well, beating \RNAblueprint{} most of the time with only a few exceptions for particularly hard instances. The results for \RNAblueprint{} are shown in Figure~\ref{fig:run-times-rbp}.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{\todo[inline]{caption Stefan?}}
  \label{fig:run-times-redprint}
\end{figure}
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{\todo[inline]{caption Stefan?}}
  \label{fig:run-times-rbp}
\end{figure}

\end{revision}
